---
title: "FIRMS Global Tuna Atlas: the IOTC nominal catch dataset, `r paste(START_YEAR, END_YEAR, sep = '-')`"
author: "IOTC Secretariat"
output: 
   html_document2:
    css: "../templates/html_template.css"
    number_sections: false
    code_folding: show
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r general_options, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, tidy = TRUE)

pn = function(number, big.mark = ",") {
  return(prettyNum(number, big.mark = big.mark))
}
```

\tableofcontents

# Summary

Nominal catches from the Indian Ocean Tuna Commission ([IOTC](https://www.iotc.org)) correspond to the annual retained catches (i.e. excluding discards) in live weight (t) by [main area](https://data.iotc.org/reference/latest/domain/admin/#IOareasMain), species, fishing gear, and fleet ([https://www.iotc.org/data/datasets](https://www.iotc.org/data/datasets)). The dataset includes nominal catches covering the period `r paste(START_YEAR, END_YEAR, sep = "-")` for (i) the 15 tuna and tuna-like species under IOTC mandate and available at species and gear level following a breakdown procedure applied to species and gear aggregates, (ii) some tuna and tuna-like species not under IOTC mandate (e.g., shortbill spearfish), and (iii) some pelagic shark and ray species commonly caught in tuna and tuna-like fisheries whose mandatory reporting varies with fishing gears (see Annex II of [Res. 15/01](https://www.iotc.org/cmm/resolution-1501-recording-catch-and-effort-data-fishing-vessels-iotc-area-competence)). Consolidated catch data for southern bluefin tuna (_Thunnus macoyii_) are available from the Commission for the Conservation of Southern Bluefin Tuna ([CCSBT](https://www.ccsbt.org)). Nominal catches for non-IOTC species are restricted to data reported at species level while data from elasmobranch species may include some species (e.g., sharks nei) and gear aggregates (e.g., gillnet operated attached to a longline). The fishing mode, derived from the geo-referenced catch dataset, is only available for the five principal market species: albacore, bigeye tuna, skipjack tuna, swordfish, and yellowfin tuna.

The document describes the processing steps taken to produce the IOTC component of the global dataset of nominal catches following the data exchange format specifications and code lists adopted by the FIRMS Global Tuna Atlas: (i) extraction from the IOTC databases, (ii) mapping with the standard code lists adopted by the FIRMS Global Tuna Atlas, and (iii) filtering on the selected list of GTA species occurring in the Indian Ocean.

# R Scripts

```{r Libraries, child = "0.1_Libraries.Rmd", echo = TRUE, results = "hide"}
```

## Loading Code Lists

```{r ReadCodelists, child = "0.2_Read_Codelists.Rmd"}
```

## Loading Mappings

```{r ReadMappings, child = "0.3_Read_Mappings.Rmd"}
```

## Loading Filterings

```{r ReadFilterings, child = "0.4_Read_Filterings.Rmd"}
```

## Extracting

```{r Extraction, child = "1.1_NC_Extraction.Rmd"}
```

## Consolidating

```{r Consolidation, child = "1.2_NC_Mapping.Rmd"}
```

## Mapping

```{r Mapping, child = "1.3_NC_Filtering.Rmd"}
```

## Filtering

```{r Filtering, child = "1.4_NC_Consolidation.Rmd"}
```

```{r DataContents, child = "1.5_NC_Contents.Rmd"}
```

## Exporting

```{r Export, child = "1.6_NC_Export.Rmd"}
```

# Results

## Processing Steps

The overall data processing resulted in the removal of `r length(unique(SPECIES_EXCLUDED))` species from the dataset, representing `r pn(round(sum(NC_MAPPED_SPECIES_GEAR_FLEET[SPECIES_CODE %in% SPECIES_EXCLUDED, CATCH])))` t of fish. The mapping to the standard code lists reduced the number of gears from `r length(unique(NC_MAPPED_SPECIES$GEAR_CODE))` to `r length(unique(NC_MAPPED_SPECIES_GEAR$gear_type))` and the fleets from `r length(unique(NC_MAPPED_SPECIES_GEAR$FLEET_CODE))` to `r length(unique(NC_MAPPED_SPECIES_GEAR_FLEET$fishing_fleet))`. The final dataset for export to the Global Tuna Atlas included `r length(unique(NC$species))` species, `r length(unique(NC$gear_type))` gears, and `r length(unique(NC$fishing_fleet))` fishing fleets.

```{r ProcessingSteps, child = "1.7_NC_Processing_Summary.Rmd"}
```

## Nominal Catch Data Excluded from the GTA

```{r DisplayNominalCatchBySpeciesExcluded, echo = FALSE}
datatable(NC_SPECIES_EXCLUDED_TABLE, rownames = FALSE, filter = "bottom") %>% formatRound("CATCH", digits = 0) %>% formatStyle("SCIENTIFIC_NAME", `font-style` = "italic")
```

## Nominal Catch Data Exported to the GTA

### Total Catch by IOTC Main Area

```{r AreaCodeList, echo = FALSE}
datatable(NC[, .(CATCH = sum(CATCH)), keyby = .(FISHING_GROUND_CODE, FISHING_GROUND, geographic_identifier)], rownames = FALSE, options = list(dom = "t")) %>% formatRound("CATCH", digits = 0)
```

### Total Catch by Species

```{r NominalCatchBySpeciesIncluded}
# Compute catch by species
NC_SPECIES = NC[, .(YEAR_INITIAL = min(year(time_start)), YEAR_FINAL = max(year(time_end)), CATCH = sum(CATCH)), keyby = .(species)]

# Add species information
NC_SPECIES_TABLE = merge(NC_SPECIES, SPECIES[, .(code, name_en, definition, taxon_family)], by.x = "species", by.y = "code")[, .(ASFIS_CODE = species, ASFIS_FAMILY = taxon_family, ASFIS_NAME = name_en, ASFIS_SCIENTIFIC_NAME = definition, YEAR_INITIAL, YEAR_FINAL, CATCH = CATCH)][order(ASFIS_FAMILY, ASFIS_CODE)]
```

```{r DisplayNominalCatchBySpecies, echo = FALSE}
datatable(NC_SPECIES_TABLE, rownames = FALSE, filter = "bottom") %>% formatRound("CATCH", digits = 0) %>% formatStyle("ASFIS_SCIENTIFIC_NAME", `font-style` = "italic")
```

### Total Catch by Gear

```{r NominalCatchByGearIncluded}
# Compute catch by gear
NC_GEAR = NC[, .(YEAR_INITIAL = min(year(time_start)), YEAR_FINAL = max(year(time_end)), CATCH = sum(CATCH)), keyby = .(gear_type)]

# Add gear information
NC_GEAR_TABLE = merge(NC_GEAR, GEARS, by.x = "gear_type", by.y = "code")[, .(ISSCFG_CODE = gear_type, ISSCFG_LABEL = label, ISSCFG_ABBREVIATION = std_abb, YEAR_INITIAL, YEAR_FINAL, CATCH)]
```

```{r DisplayNominalCatchByGear, echo = FALSE}
datatable(NC_GEAR_TABLE, rownames = FALSE, filter = "bottom") %>% formatRound("CATCH", digits = 0)
```

### Total Catch by Fleet

```{r NominalCatchByFleetIncluded}
# Compute catch by fleet
NC_FLEET = NC[, .(YEAR_INITIAL = min(year(time_start)), YEAR_FINAL = max(year(time_end)), CATCH = sum(CATCH)), keyby = .(fishing_fleet)]

# Add gear information
NC_FLEET_TABLE = merge(NC_FLEET, FLEETS, by.x = "fishing_fleet", by.y = "code")[, .(FIRMS_FLEET_CODE = fishing_fleet, FIRMS_FLEET = label, YEAR_INITIAL, YEAR_FINAL, CATCH)]
```

```{r DisplayNominalCatchByFleet, echo = FALSE}
datatable(NC_FLEET_TABLE, rownames = FALSE, filter = "bottom") %>% formatRound("CATCH", digits = 0)
```

# Data Exchange Format

```{r TableExportFormat, echo = FALSE}
datatable(DDD, rownames = FALSE, options = list(dom = "t", pageLength = 14)) %>% 
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Administrative", "#EFF3FF")) %>%
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Temporal", "#BDD7E7")) %>%
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Spatial", "#6BAED6")) %>%
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Technical", "#3182BD")) %>%
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Measurement", "#08519C")) %>% 
  formatStyle(columns = "ENTITY", target = "row", color = styleEqual("Measurement", "white"))
```

# Reference Code Lists

## Species

```{r SpeciesCodeList, echo = FALSE}
datatable(SPECIES_CODE_LIST, rownames = FALSE, filter = "bottom", options = list(dom = "Bfrtip", buttons = list(list(extend = "csv")))) %>% formatStyle("definition", `font-style` = "italic")
```

## Gears

```{r GearCodeList, echo = FALSE}
datatable(GEAR_CODE_LIST, rownames = FALSE, filter = "bottom",  options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))
```

## Fleets

```{r FleetCodeList, echo = FALSE}
datatable(FLEET_CODE_LIST, rownames = FALSE, filter = "bottom", extensions = "Buttons", options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))
```

# Mapping Tables

## IOTC Species Codes to ASFIS

```{r MappingSpeciesTable, echo = FALSE}
datatable(MAPPING_SPECIES[src_code != trg_code, .(IOTC_SPECIES_CODE = src_code, ASFIS_CODE = trg_code, ASFIS_NAME = name_en, ASFIS_SCIENTIFIC_NAME = definition)][order(IOTC_SPECIES_CODE)], rownames = FALSE, filter = "bottom",  options = list(dom = "Bfrtip", buttons = list(list(extend = "csv")))) %>% formatStyle("ASFIS_SCIENTIFIC_NAME", `font-style` = "italic")
```

## IOTC Gear Codes to ISSCFG

```{r MappingGearsTable, echo = FALSE}
datatable(MAPPING_GEAR[, .(IOTC_GEAR_CODE = src_code, ISSCFG_CODE = trg_code, ISSCFG_LABEL = label, ISSCFG_ABBREVIATION = std_abb)][order(IOTC_GEAR_CODE)], rownames = FALSE, filter = "bottom",  options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))
```

## IOTC Fleet Codes to FIRMS Fishing Fleets

```{r MappingFleetsTable, echo = FALSE}
datatable(MAPPING_FLEET[src_code != trg_code, .(IOTC_FLEET_CODE = src_code, FIRMS_FLEET_CODE = trg_code, FIRMS_FLEET_LABEL = label)][order(IOTC_FLEET_CODE)], rownames = FALSE, filter = "bottom",  options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))
```
