---
title: "FIRMS Global Tuna Atlas: the IOTC geo-referenced catch dataset, 1952-`r END_YEAR`"
author: "IOTC Secretariat"
output: 
   html_document2:
    css: "../templates/html_template.css"
    number_sections: FALSE
    code_folding: show
    toc: true
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
---
  
```{r general_options, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, tidy = TRUE)

pn = function(number, big.mark = ",") {
  return(prettyNum(number, big.mark = big.mark))
}
```

# Summary

Geo-referenced catch data correspond to the retained catch in live weight (t) or numbers, by species, fishing gear, grid area, month, and fleet ([https://www.iotc.org/data/datasets](https://www.iotc.org/data/datasets)). The IOTC geo-referenced catch dataset includes all catches from all coastal and offshore baitboat, gillnet, line, longline, purse seine, and other gears (e.g., lift nets) fisheries available by month and grid area covering the period 1952-2021 for (i) the 15 tuna and tuna-like species under IOTC mandate, (ii) some tuna and tuna-like species not under IOTC mandate (e.g., shortbill spearfish), (iii) some pelagic shark and ray species commonly caught in tuna and tuna-like fisheries whose reporting to the IOTC varies with fishing gears (see Annex II of [Res. 15/01](https://www.iotc.org/cmm/resolution-1501-recording-catch-and-effort-data-fishing-vessels-iotc-area-competence)), and (iv) some species aggregates that include some of the species previously mentioned, e.g., true tunas nei. Consolidated catch data for southern bluefin tuna (_Thunnus macoyii_) are available from the Commission for the Conservation of Southern Bluefin Tuna ([CCSBT](https://www.ccsbt.org)). 

For fishing vessels recorded in the IOTC Record of Authorized Vessels ([IOTC Res. 19/04](https://www.iotc.org/cmm/resolution-1904-concerning-iotc-record-vessels-authorised-operate-iotc-area-competence)), geo-referenced catch data should be extrapolated to the annual catch and reported on a monthly basis by 1°x1° and 5°x5° grid area for surface and longline fisheries, respectively ([IOTC Res. 15/02](https://www.iotc.org/cmm/resolution-1502-mandatory-statistical-reporting-requirements-iotc-contracting-parties-and)). In the case of coastal (i.e., artisanal) fisheries, catch data may be reported by grid area or using alternative geographical areas that were approximated by regular 1°x1° or 5°x5° grid areas when possible.

The document describes the processing steps taken to produce the IOTC component of the global geo-referenced dataset of fisheries catches: (i) data extraction from the IOTC databases, (ii) mapping with the standard code lists adopted by the FIRMS Global Tuna Atlas, and (iii) filtering on the selected list of GTA species occurring in the Indian Ocean and strata of spatial resolution of 1°x1° or 5°x5° and temporal resolution of 1 month.

# R Scripts

```{r Libraries, child = "0.1_Libraries.Rmd", echo = TRUE, results = "hide"}
```

## Loading Code Lists

```{r ReadCodelists, child = "0.2_Read_Codelists.Rmd"}
```

## Loading Mappings

```{r ReadMappings, child = "0.3_Read_Mappings.Rmd"}
```

## Loading Filterings

```{r ReadFilterings, child = "0.4_Read_Filterings.Rmd"}
```

## Loading Spatial Layers

```{r ReadSpatialLayers, child = "0.5_Read_Spatial_Layers.Rmd"}
```

## Extracting

```{r Extraction, child = "2.1_CA_Extraction.Rmd"}
```

## Consolidating

```{r Consolidation, child = "2.2_CA_Consolidation.Rmd"}
```

## Mapping

```{r Mapping, child = "2.3_CA_Mapping.Rmd"}
```

## Filtering

```{r Filtering, child = "2.4_CA_Filtering.Rmd"}
```

```{r DataContents, child = "2.5_CA_Contents.Rmd"}
```

## Exporting

```{r Export, child = "2.6_CA_Export.Rmd"}
```

# Results

The data processing resulted in the removal of `r pn(round(sum(CA_M_MAPPED_SPECIES_GEAR_FLEET[CATCH_UNIT_CODE == "t", CATCH]) - sum(CA[CATCH_UNIT_CODE == "t", CATCH])))` t of fish and `r pn(round(sum(CA_M_MAPPED_SPECIES_GEAR_FLEET[CATCH_UNIT_CODE == "no", CATCH]) - sum(CA[CATCH_UNIT_CODE == "no", CATCH])))` fish from the original dataset through the filtering of species and spatio-temporal strata. The mapping to the standard code lists reduced the number of gears from `r length(unique(CA_M_MAPPED_SPECIES$GEAR_CODE))` to `r length(unique(CA_M_MAPPED_SPECIES_GEAR$gear_type))` and the fleets from `r length(unique(CA_M_MAPPED_SPECIES$FLEET_CODE))` to `r length(unique(CA_M_MAPPED_SPECIES_GEAR_FLEET$fishing_fleet))`. The filtering on the GTA species occurrin gin the Indian Ocean  removed a total of `r length(unique(CA_M_MAPPED_SPECIES_GEAR_FLEET$species)) - length(unique(CA$species))` taxa.

## Processing Steps

```{r ProcessingSteps, child = "2.7_CA_Processing_Summary.Rmd"}
```

## Geo-Referenced Catch Data Excluded from the GTA

The overall data processing resulted in the removal of a total of `r pn(round( sum(CA_RAW[unit == "t", CATCH]) - sum(CA[unit == "t", CATCH])))` t and `r pn(round( sum(CA_RAW[unit == "no", CATCH]) - sum(CA[unit == "no", CATCH])))` from the initial dataset. The filtering removed `r length(unique(SPECIES_EXCLUDED))` species from the dataset, representing `r pn(round(sum(CA_SPECIES_EXCLUDED_TABLE[UNIT == "t", CATCH])))` t and `r pn(round(sum(CA_SPECIES_EXCLUDED_TABLE[UNIT == "no", CATCH])))` fish. In addition, a total of `r length(unique(CA_NON_REGULAR_GRID$FISHING_GROUND_CODE))` spatial grids of resolution larger than 5°x5° were removed from the dataset, representing `r pn(round(sum(CA_NON_REGULAR_GRID[CATCH_UNIT_CODE == "t", CATCH])))` t of fish. Furthermore, a catch of `r pn(round(CA_REGULAR_GRIDS_KO_TABLE[UNIT == "no", sum(CATCH)]))` fish and `r pn(round(CA_REGULAR_GRIDS_KO_TABLE[UNIT == "t", sum(CATCH)]))` t of fish was removed from the dataset as it was reported for `r length(unique(CA_REGULAR_GRIDS_KO$FISHING_GROUND_CODE))` regular grids located outside the Indian Ocean. Also, some catch data were reported with a temporal resolution larger than 1 month, corresponding to a total of `r pn(round(sum(CA_EXCLUDED_TEMPORAL[CATCH_UNIT_CODE == "t", CATCH])))` t and `r pn(round(sum(CA_EXCLUDED_TEMPORAL[CATCH_UNIT_CODE == "no", CATCH])))` fish removed from the original dataset.

### Species

```{r DisplayCatchBySpeciesExcluded}
datatable(CA_SPECIES_EXCLUDED_TABLE, rownames = FALSE, options = list(dom = "t")) %>% formatRound("CATCH", digits = 1) %>% formatStyle("SCIENTIFIC_NAME", `font-style` = "italic")
```

### Spatial Strata 

```{r DisplaySpatialStrataExcluded}
datatable(CA_EXCLUDED_SPATIAL_TABLE, rownames = FALSE, options = list(dom = "t")) %>% formatRound("CATCH", digits = 0)
```

### Temporal Strata

```{r DisplayTemporalStrataExcluded}
datatable(CA_EXCLUDED_TEMPORAL_TABLE, rownames = FALSE, options = list(dom = "t")) %>% formatRound("CATCH", digits = 0)
```

## Geo-Referenced Catch Data Exported to the GTA

### Total Catch by Species

```{r CatchBySpeciesIncluded, results = "hide", include = FALSE, message = FALSE}
# Compute catch by species
CA_SPECIES = CA[, .(YEAR_INITIAL = min(year(time_start)), YEAR_FINAL = max(year(time_end)), CATCH = sum(CATCH)), keyby = .(species, UNIT = CATCH_UNIT_CODE)]

# Add species information
CA_SPECIES_TABLE = merge(CA_SPECIES, SPECIES[, .(code, name_en, definition, taxon_family)], by.x = "species", by.y = "code", all.x = TRUE)[, .(CODE = species, FAMILY = taxon_family, NAME = name_en, SCIENTIFIC_NAME = definition, YEAR_INITIAL, YEAR_FINAL, UNIT, CATCH = CATCH)][order(FAMILY, CODE)]

```

```{r DisplayCatchBySpecies}
datatable(CA_SPECIES_TABLE, rownames = FALSE, filter = "bottom") %>% formatRound("CATCH", digits = 1) %>% formatStyle("CATCH", `text-align` = "right") %>% formatStyle("SCIENTIFIC_NAME", `font-style` = "italic")
```

### Total Catch by Gear

```{r CatchByGearIncluded, results = "hide", include = FALSE, message = FALSE}
# Compute catch by gear
CA_GEAR = CA_M_MAPPED_SPECIES_GEAR_FLEET[, .(YEAR_INITIAL = min(year(time_start)), YEAR_FINAL = max(year(time_end)), CATCH = sum(CATCH)), keyby = .(gear_type, UNIT = CATCH_UNIT_CODE)]

# Add gear information
CA_GEAR_TABLE = merge(CA_GEAR, GEARS, by.x = "gear_type", by.y = "code")[, .(ISSCFG_CODE = gear_type, ISSCFG_LABEL = label, ISSCFG_ABBREVIATION = std_abb, YEAR_INITIAL, YEAR_FINAL, UNIT, CATCH)]
```

```{r DisplayCatchByGear}
datatable(CA_GEAR_TABLE, rownames = FALSE, filter = "bottom") %>% formatRound("CATCH", digits = 0)
```

### Total Catch by Fleet

```{r CatchByFleetIncluded, results = "hide", include = FALSE, message = FALSE}
# Compute catch by fleet
CA_FLEET = CA_M_MAPPED_SPECIES_GEAR_FLEET[, .(YEAR_INITIAL = min(year(time_start)), YEAR_FINAL = max(year(time_end)), CATCH = sum(CATCH)), keyby = .(fishing_fleet, UNIT = CATCH_UNIT_CODE)]

# Add gear information
CA_FLEET_TABLE = merge(CA_FLEET, FLEETS, by.x = "fishing_fleet", by.y = "code")[, .(FIRMS_FLEET_CODE = fishing_fleet, FIRMS_FLEET = label, YEAR_INITIAL, YEAR_FINAL, UNIT, CATCH)]
```

```{r DisplayCatchByFleet}
datatable(CA_FLEET_TABLE, rownames = FALSE, filter = "bottom") %>% formatRound("CATCH", digits = 0) 
```

# Data Exchange Format

```{r TableExportFormat}
datatable(DDD, rownames = FALSE, options = list(dom = "t", pageLength = 14)) %>% 
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Administrative", "#EFF3FF")) %>%
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Temporal", "#BDD7E7")) %>%
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Spatial", "#6BAED6")) %>%
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Technical", "#3182BD")) %>%
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Measurement", "#08519C")) %>% 
  formatStyle(columns = "ENTITY", target = "row", color = styleEqual("Measurement", "white"))
```

# Reference Code Lists

## Species

```{r SpeciesCodeList}
datatable(SPECIES_CODE_LIST, rownames = FALSE, filter = "bottom", options = list(dom = "Bfrtip", buttons = list(list(extend = "csv")))) %>% formatStyle("definition", `font-style` = "italic")
```

## Gears

```{r GearCodeList}
datatable(GEAR_CODE_LIST, rownames = FALSE, filter = "bottom",  options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))
```

## Fleets

```{r FleetCodeList}
datatable(FLEET_CODE_LIST, rownames = FALSE, filter = "bottom", extensions = "Buttons", options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))
```

# Mapping Tables

## IOTC Species Codes to ASFIS

```{r MappingSpeciesTable}
datatable(MAPPING_SPECIES[src_code != trg_code, .(IOTC_SPECIES_CODE = src_code, ASFIS_CODE = trg_code, ASFIS_NAME = name_en, ASFIS_SCIENTIFIC_NAME = definition)][order(IOTC_SPECIES_CODE)], rownames = FALSE, filter = "bottom",  options = list(dom = "Bfrtip", buttons = list(list(extend = "csv")))) %>% formatStyle("ASFIS_SCIENTIFIC_NAME", `font-style` = "italic")
```

## IOTC Gear Codes to ISSCFG

```{r MappingGearsTable}
datatable(MAPPING_GEAR[, .(IOTC_GEAR_CODE = src_code, ISSCFG_CODE = trg_code, ISSCFG_LABEL = label, ISSCFG_ABBREVIATION = std_abb)][order(IOTC_GEAR_CODE)], rownames = FALSE, filter = "bottom",  options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))

```

## IOTC Fleet Codes to FIRMS Fishing Fleets

```{r MappingFleetsTable}
datatable(MAPPING_FLEET[src_code != trg_code, .(IOTC_FLEET_CODE = src_code, FIRMS_FLEET_CODE = trg_code, FIRMS_FLEET_LABEL = label)][order(IOTC_FLEET_CODE)], rownames = FALSE, filter = "bottom",  options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))
```

