---
title: "FIRMS Global Tuna Atlas: the IOTC geo-referenced catch data set by month and 5°x5° grid area, 1952-2021"
author: "IOTC Secretariat"
output: 
   html_document2:
    css: "../templates/html_template.css"
    number_sections: FALSE
    code_folding: show
    toc: true
    toc_float:
      collapsed: FALSE
      smooth_scroll: FALSE
---
  
```{r general_options, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE, tidy = TRUE)

pn = function(number, big.mark = ",") {
  return(prettyNum(number, big.mark = big.mark))
}
```

# SUMMARY

Geo-referenced catch data correspond to the catch in live weight (t) or numbers, by species, fishing gear, grid area, month, and fleet ([https://www.iotc.org/data/datasets](https://www.iotc.org/data/datasets)). The IOTC geo-referenced catch data set by month and 5°x5° grid area includes all catches from all coastal and offshore baitboat, gillnet, line, longline, purse seine, and other gears (e.g., lift nets) fisheries available by month and 5°x5° grid area covering the period 1952-2021 for (i) the 16 tuna and tuna-like species under IOTC mandate, (ii) some tuna and tuna-like species not under IOTC mandate (e.g., shortbill spearfish), (iii) some pelagic shark and ray species commonly caught in tuna and tuna-like fisheries whose reporting to the IOTC varies with fishing gears (see Annex II of [Res. 15/01](https://www.iotc.org/cmm/resolution-1501-recording-catch-and-effort-data-fishing-vessels-iotc-area-competence)), and (iv) some species aggregates that include some of the species previously mentioned, e.g., true tunas nei.

For fishing vessels recorded in the IOTC Record of Authorized Vessels ([IOTC Res. 19/04](https://www.iotc.org/cmm/resolution-1904-concerning-iotc-record-vessels-authorised-operate-iotc-area-competence)), geo-referenced catch data should be extrapolated to the annual catch and reported on a monthly basis by 1°x1° and 5°x5° grid area for surface and longline fisheries, respectively ([IOTC Res. 15/02](https://www.iotc.org/cmm/resolution-1502-mandatory-statistical-reporting-requirements-iotc-contracting-parties-and)). In the case of coastal (i.e., artisanal) fisheries, catch data may be reported by grid area or using alternative geographical areas that were approximated by regular 5°x5° grid areas when possible.

The document describes the processing steps taken to produce the IOTC component of the global data set of fisheries catches by 5°x5° grid area and month strata: (i) data extraction from the IOTC databases, (ii) filtering to remove species not occurring in the Indian Ocean, strata of spatial resolution larger than 5°x5° and temporal resolution longer than 1 month, and (iii) mapping with the standard code lists adopted by the FIRMS Global Tuna Atlas.

# R SCRIPTS

```{r Libraries, child = "0_1.Libraries.Rmd", echo = TRUE, results = "hide"}
```

## Loading code lists and mapping

```{r CodeLists, child = "0_2.Code_Lists_Mapping.Rmd"}
```

## Extracting

```{r Extraction, child = "3_1.CA_5deg_1m_Extraction.Rmd"}
```

## Filtering

```{r Filtering, child = "3_2.CA_5deg_1m_Filtering.Rmd"}
```

## Consolidating

```{r Consolidation, child = "3_3.CA_5deg_1m_Consolidation.Rmd"}
```

## Mapping

```{r Mapping, child = "3_4.CA_5deg_1m_Mapping.Rmd"}
```

## Exporting

```{r Export, child = "3_5.CA_5deg_1m_Export.Rmd"}
```

# RESULTS

The data processing resulted in the removal of `r pn(round(sum(CA_RAW[CATCH_UNIT_CODE == "t", CATCH]) - sum(CA_55M_MAPPED_SPECIES_GEAR_FLEET[CATCH_UNIT_CODE == "t", CATCH])))` t of fish and `r pn(round(sum(CA_RAW[CATCH_UNIT_CODE == "no", CATCH]) - sum(CA_55M_MAPPED_SPECIES_GEAR_FLEET[CATCH_UNIT_CODE == "no", CATCH])))` fish from the original data set through the filtering of species and spatio-temporal strata. The mapping to the standard code lists reduced the number of gears from `r length(unique(CA_55M_MAPPED_SPECIES$GEAR_CODE))` to `r length(unique(CA_55M_MAPPED_SPECIES_GEAR$gear))` and the fleets from `r length(unique(CA_55M_MAPPED_SPECIES$FLEET_CODE))` to `r length(unique(CA_55M_MAPPED_SPECIES_GEAR_FLEET$fleet))`.

## Processing steps

```{r ProcessingSteps, child = "3_6.CA_5deg_1m_Processing_Summary.Rmd"}
```

## Geo-referenced catch data excluded from the GTA

The overall data processing resulted in the removal of `r length(unique(SPECIES_EXCLUDED))` species from the data set, representing `r pn(round(sum(CA_SPECIES_EXCLUDED_TABLE[UNIT == "t", CATCH])))` t of fish and `r pn(round(sum(CA_SPECIES_EXCLUDED_TABLE[UNIT == "no", CATCH])))` fish. A total of `r length(unique(CA_EXCLUDED_SPATIAL$FISHING_GROUND_CODE))` spatial grids of resolution larger than 5°x5° were removed from the data set, representing `r pn(round(sum(CA_EXCLUDED_SPATIAL[CATCH_UNIT_CODE == "t", CATCH])))` t of fish. Furthermore, some catch data were reported with a temporal resolution larger than 1 month, corresponding to a total of `r pn(round(sum(CA_EXCLUDED_TEMPORAL[CATCH_UNIT_CODE == "t", CATCH])))` t of fish and `r pn(round(sum(CA_EXCLUDED_TEMPORAL[CATCH_UNIT_CODE == "no", CATCH])))` fish removed from the original data set.

### Species not occurring in the Indian Ocean

```{r DisplayCatchBySpeciesExcluded, echo = FALSE}
datatable(CA_SPECIES_EXCLUDED_TABLE, rownames = FALSE, options = list(dom = "t")) %>% formatRound("CATCH", digits = 1) %>% formatStyle("ASFIS_SCIENTIFIC_NAME", `font-style` = "italic")
```

### Strata of spatial resolution larger than 5°x5°

```{r DisplaySpatialStrataExcluded}
datatable(CA_EXCLUDED_SPATIAL_TABLE, rownames = FALSE, options = list(dom = "t")) %>% formatRound("CATCH", digits = 0)
```

### Strata of temporal resolution larger than 1 month

```{r DisplayTemporalStrataExcluded}
datatable(CA_EXCLUDED_TEMPORAL_TABLE, rownames = FALSE, options = list(dom = "t")) %>% formatRound("CATCH", digits = 0)
```

## Geo-referenced catch data exported to the GTA

### Total catch by species

```{r CatchBySpeciesIncluded}
# Compute catch by species
CA_SPECIES = CA_55M_MAPPED_SPECIES_GEAR_FLEET[, .(YEAR_INITIAL = min(year(time_start)), YEAR_FINAL = max(year(time_end)), CATCH = sum(CATCH)), keyby = .(species, UNIT = CATCH_UNIT_CODE)]

# Add species information
CA_SPECIES_TABLE = merge(CA_SPECIES, SPECIES[, .(`3A_CODE`, English_name, Scientific_name, Family)], by.x = "species", by.y = "3A_CODE")[, .(ASFIS_CODE = species, ASFIS_FAMILY = Family, ASFIS_NAME = English_name, ASFIS_SCIENTIFIC_NAME = Scientific_name, YEAR_INITIAL, YEAR_FINAL, UNIT, CATCH)][order(ASFIS_FAMILY, ASFIS_CODE)]
```

```{r DisplayCatchBySpecies, echo = FALSE}
datatable(CA_SPECIES_TABLE, rownames = FALSE, filter = "bottom") %>% formatRound("CATCH", digits = 0) %>% formatStyle("CATCH", `text-align` = "right") %>% formatStyle("ASFIS_SCIENTIFIC_NAME", `font-style` = "italic")
```

### Total catch by gear

```{r CatchByGearIncluded, echo = FALSE}
# Compute catch by gear
CA_GEAR = CA_55M_MAPPED_SPECIES_GEAR_FLEET[, .(YEAR_INITIAL = min(year(time_start)), YEAR_FINAL = max(year(time_end)), CATCH = sum(CATCH)), keyby = .(gear, UNIT = CATCH_UNIT_CODE)]

# Add gear information
CA_GEAR_TABLE = merge(CA_GEAR, GEARS, by.x = "gear", by.y = "code")[, .(ISSCFG_GEAR_CODE = gear, ISSCFG_GEAR = label, ISSCFG_ABB = standard_abbreviations, YEAR_INITIAL, YEAR_FINAL, UNIT, CATCH)]
```

```{r DisplayCatchByGear, echo = FALSE}
datatable(CA_GEAR_TABLE, rownames = FALSE, filter = "bottom") %>% formatRound("CATCH", digits = 0)
```

### Total catch by fleet

```{r CatchByFleetIncluded, echo = FALSE}
# Compute catch by fleet
CA_FLEET = CA_55M_MAPPED_SPECIES_GEAR_FLEET[, .(YEAR_INITIAL = min(year(time_start)), YEAR_FINAL = max(year(time_end)), CATCH = sum(CATCH)), keyby = .(fleet, UNIT = CATCH_UNIT_CODE)]

# Add gear information
CA_FLEET_TABLE = merge(CA_FLEET, FLEETS, by.x = "fleet", by.y = "code")[, .(FIRMS_FLEET_CODE = fleet, FIRMS_FLEET = label, YEAR_INITIAL, YEAR_FINAL, UNIT, CATCH)]
```

```{r DisplayCatchByFleet, echo = FALSE}
datatable(CA_FLEET_TABLE, rownames = FALSE, filter = "bottom") %>% formatRound("CATCH", digits = 0) 
```

# DATA EXCHANGE FORMAT

```{r TableExportFormat, echo = FALSE}
datatable(DDD, rownames = FALSE, options = list(dom = "t", pageLength = 14)) %>% 
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Administrative", "#EFF3FF")) %>%
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Temporal", "#BDD7E7")) %>%
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Spatial", "#6BAED6")) %>%
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Technical", "#3182BD")) %>%
  formatStyle(columns = "ENTITY", target = "row", backgroundColor = styleEqual("Measurement", "#08519C")) %>% 
  formatStyle(columns = "ENTITY", target = "row", color = styleEqual("Measurement", "white"))
```

# REFERENCE CODE LISTS

## Species

```{r SpeciesCodeList, echo = FALSE}
datatable(SPECIES_CODE_LIST, rownames = FALSE, filter = "bottom", options = list(dom = "Bfrtip", buttons = list(list(extend = "csv")))) %>% formatStyle("Scientific_name", `font-style` = "italic")
```

## Gears

```{r GearCodeList, echo = FALSE}
datatable(GEAR_CODE_LIST, rownames = FALSE, filter = "bottom",  options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))
```

## Fleets

```{r FleetCodeList, echo = FALSE}
datatable(FLEET_CODE_LIST, rownames = FALSE, filter = "bottom", extensions = "Buttons", options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))
```

# MAPPING TABLES

## IOTC species codes to ASFIS

```{r MappingSpeciesTable, echo = FALSE}
datatable(MAPPING_SPECIES[src_code != trg_code, .(IOTC = src_code, ASFIS = trg_code, `ASFIS scientific name` = Scientific_name, `ASFIS English name` = English_name)][order(IOTC)], rownames = FALSE, filter = "bottom", options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))  %>% formatStyle("ASFIS scientific name", `font-style` = "italic")
```

## IOTC gear codes to ISSCFG

```{r MappingGearsTable, echo = FALSE}
datatable(MAPPING_GEAR[, .(IOTC = src_code, ISSCFG = trg_code, `ISSCFG abbreviation` = standard_abbreviations, `ISSCFG label` = label)][order(IOTC)], rownames = FALSE, filter = "bottom", options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))
```

## IOTC fleet codes to FIRMS fishing fleets

```{r MappingFleetsTable, echo = FALSE}
datatable(MAPPING_FLEET[src_code != trg_code, .(IOTC = src_code, FIRMS = trg_code, `FIRMS label` = label)][order(IOTC)], rownames = FALSE, filter = "bottom", options = list(dom = "Bfrtip", buttons = list(list(extend = "csv"))))
```

